A Methodology and Implementation of Automated Emissions Harmonization for Use in Integrated Assessment Models

### 2017-05-24 intro.tex
intro edited

Integrated Assessment Models (IAMs) are tools used to understand the complex
interactions between energy systems, economoic systems, land use, the climate,
and air pollution, among other sectors \TODO{cite some IAMs}. IAMs provide
global projections of systemic change by dividing the world into a number of
representative regions, the definition of which is distinct for each
model. Results from IAMs are integral in a number of international studies,
which notably include projections of climate and economic futures used in
analyses underpinning the Intergovenmental Panel on Climate Change (IPCC)
\TODO{cite} and the United Nations Framework Convention on Climate Unchange
(UNFCCC)'s Paris Agreement \TODO{cite}.

@@

Future emissions pathways are of critical importance to both of global and local
governance concerns. For example, the United Nation's Sustainable Development
Goals (SDGs) \TODO{CITE} speak directly to the impacts of emissions in Goal 3
(Good Health and Well-Being) and Goal 13 (Climate Action). IAMs endogenously
calculate many of the emissions species of import that impact the progress
towards these SDGs by modeling the individual technologies and sectors that
contribute towards the emissions of different air pollutants and GHGs as well as
various mitigation technologies (e.g., NOx and SOx scrubbers on power plants)
that are used due to regulation or implemented policies.

In practice, IAMs calculate the total source intensity of emitting technologies,
for example the total activity of coal power plant emissions in China, and
incorporate emissions-intensity factors for individual species, for example the
quantity of sulfur emissions from coal plants per megawatt-hour of production,
in order to determine total emissions of a given gas species from a given
technology. Models are then \textit{calibrated} to historical data sources which
provide both the activity of emitting technologies as well as trajectories of
past emissions in order to determine historical emissions factors in each model
region for each technology.

With some frequency, however, historical data is updated to incorporate new
assumptions of source activity, new emissions accounting practices, or other
concerns. The global climate change community has recently updated a standard
historical emissions dataset for both anthropogenic and land-use change
emissions \TODO{cite ceds and LUC paper} as part of the 6th phase of the Coupled
Model Intercomparison Project \TODO{cite cmip6} (CMIP6). These historical
emissions trajectories in conjunction with future trajectories as provided by
IAMs are critical inputs to the climate-related modeling exercises of CMIP6.

Because the process of model calibration is quite onerous, model teams initially
incorporate updated historical datasets through the process of
\textit{harmonization} in order to guarantee consistency with new historical
datasets. Harmonization describes the process of updating model results to
incorporate new historical timeseries while maintaining critical aspects of
modeled future trajectories for variables of import. In the emissions context,
this means that each individual combination of model region, model sector, and
emissions species must be harmonized. Depending on the model, this can require
the selection of thousands to tens-of-thousands of harmonization methods.

Harmonization has been addressed in previous studies as it is a common practice
in the IAM and climate change communities. For example, \TODO{cite Malte}
describes a harmonization process using scaling routines for the 5 SRES regions
with no sectoral definition. Further, \TODO{cite Joeri} describes the impacts of
choosing various harmonization routines on future trajectories. Importantly, the
choice of harmonization method to date has been determined via expert
solicitation and has generally been applied to all trajectories for a given
class of emissions species.

In order to incoporate the variety of possible model results and given
relatively large increases in regional and sectoral dimensionality, an automated
process for determining harmonization methods is needed. The remainder of this
paper describes the methodology of the \code{aneris} codebase which implements
such a system in Section \ref{sec:meths}. The results of applying the automated
harmonization mechanism on two example IAM scenarios, one with emissions growth
and another with emissions mitigation, is presented in Section
\ref{sec:results}. Finally, the general effectiveness and potential future
improvements on the automated methodology is discussed in Section
\ref{sec:future}.

>>>>

Integrated Assessment Models (IAMs) are tools used to understand the complex
interactions between energy systems, economic systems, land use, the climate,
and air pollution, among other sectors . IAMs provide global projections of
systemic change by dividing the world into a number of representative regions,
the definition of which is distinct for each model. Results from IAMs are
integral in a number of international studies, which notably include projections
of climate and economic futures. Recently, flagship IAMs have developed
scenarios based on the Shared Socioeconomic Pathways (SSPs)
\cite{van_vuuren_energy_2017, kriegler_fossil-fueled_2017, calvin_ssp4:_2017,
  fujimori_ssp3:_2017, fricko_marker_2017} which model a variety of potential
global futures. The SSPs are used in analyses underpinning the Intergovernmental
Panel on Climate Change (IPCC) and the United Nations Framework Convention on
Climate Change (UNFCCC)'s Paris Agreement \TODO{cite}.

@@

Future emissions pathways are of critical importance to both of global and local
governance concerns. For example, the United Nation's Sustainable Development
Goals (SDGs) \TODO{CITE} speak directly to the impacts of emissions in Goal 3
(Good Health and Well-Being) and Goal 13 (Climate Action). IAMs endogenously
calculate the trajectories many of the emissions species of import that impact
the progress towards these SDGs by modeling the individual technologies and
sectors that contribute towards the emissions of different air pollutants and
GHGs as well as various mitigation technologies (e.g., NOx and SOx scrubbers on
power plants).

In practice, IAMs calculate the total source intensity of emitting technologies,
for example the total activity of coal power plants in China, and incorporate
emissions-intensity factors for individual gas species, for example the quantity
of sulfur emissions from coal plants per megawatt-hour of production, in order
to determine total emissions of a given gas species from a given
technology. Models are then \textit{calibrated} to historical data sources which
provide both the activity of emitting technologies as well as trajectories of
past emissions in order to determine historical emissions factors in each model
region for each technology.

With some frequency, however, historical data is updated to incorporate new
assumptions of source activity, new emissions accounting practices, or other
concerns. The global climate change community has recently developed a standard
historical emissions dataset for both anthropogenic
\cite{hoesly_historical_2017} and land-use change emissions
\cite{van_marle_historic_2017} as part of the 6th phase of the Coupled Model
Inter-comparison Project (CMIP6). These historical emissions trajectories in
conjunction with future trajectories as provided by IAMs are critical inputs to
the climate-related modeling exercises of CMIP6.

Because the process of model calibration is quite onerous, model teams initially
incorporate updated historical datasets through the process of
\textit{harmonization} in order to guarantee consistency with these new
historical datasets. Harmonization describes the process of updating model
results to incorporate new historical time series while maintaining critical
aspects of modeled future trajectories impacted outputs. In the emissions
context, this means that each individual combination of model region, model
sector, and emissions species must be harmonized. Depending on the total number
of model regions, sectors, and emissions species, this can require the selection
of thousands to tens-of-thousands of harmonization methods.

Harmonization has been addressed in previous studies as it is a common practice
in the IAM and climate change communities. For example,
\cite{meinshausen_rcp_2011} describes the use of scaling routines for the 5
regions used in the Special Report on Emissions Scenarios \TODO{CITE}; however,
only total emissions were harmonized in the exercise, thus there is no sectoral
dimension. Further, \cite{rogelj_discrepancies_2011} describes the impacts of
choosing various harmonization routines on future trajectories. Importantly, the
choice of harmonization method to date has been determined via expert
solicitation and has generally been applied to all trajectories for a given
class of emissions species.

In order to incorporate the large variety of possible model results and given
relatively large increases in regional and sectoral dimensionality, an automated
process for determining harmonization methods is needed. The remainder of this
paper describes the methodology of the \code{aneris} code base, written in the
Python programming language, which implements such a system in Section
\ref{sec:meths}. The results of applying the automated harmonization mechanism
on two example IAM scenarios, one with emissions growth and another with
emissions mitigation, is presented in Section \ref{sec:results}. Finally, the
general effectiveness and potential future improvements on the automated
methodology is discussed in Section \ref{sec:future}.
